---
title: "week16_recap"
author: "Daniel Trielli"
date: "2023-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(tigris)
library(tidycensus)
library(corrr)
library(tidytext)
```

## Load Data

[USAFacts compiled data on inspections on child care facilities throughout the country](https://usafacts.org/data-projects/childcare-inspections) - specifically, how many of those facilities are overdue for inspection. Let's take a look on the data they have on Maryland.

First, here the data we need to load:

```{r}
childcare_inspections <- read_csv('data/childcare_inspections.csv')
childcare_inspections$fips <- as.character(childcare_inspections$fips)
```

#### **Q1** The childcare data does not have the names of the counties. Join them with the md_counties data, that can be loaded in the block below. Then inspect the dataframe you created with this join and answer this: there are two columns with county names in the new dataframe. Which one is the best to be used? (hint: it has to do with homonyms)

```{r}
# This loads the county data
md_counties <- counties() |>  filter(STATEFP == "24")
```

```{r}

childcare_with_counties <- childcare_inspections |> 
left_join(md_counties, join_by (fips ==GEOID))
# Join the childcare data with the county data 

```

##### A1: Countyfp would be the better one since it is a universal indicator and would work best for combining and merging the datasets. 

------------------------------------------------------------------------

#### **Q2** Create bar chart with the percentages of overdue childcare facility inspections per county to answer (You'll need to create a column for that calculation first). What is the county with the largest percentage of facilities with overdue inspections?

```{r}
childcare_with_counties <-childcare_with_counties |> 
  mutate(percent_overdue = (overdue/total) * 100) 

# Calculate the percentage of overdue inspections per county

# Create the chart with ggplot
  childcare_with_counties |>
  ggplot(aes(x = NAME, y = percent_overdue)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Overdue Childcare Inspections Per County in MD",
    x = "County",
    y = "Percentage of Overdue Inspections",
    caption = "Source: USAFacts"
  )
```

##### **A2: Prince Geoge's is the county with the highest percentage of overdue inspections.** 

------------------------------------------------------------------------

#### **Q3** Next, we want to see if the number of child care facilities - with overdue inspections or otherwise - is appropriate for each county. So we're going to calculate the rate of children per facility per county. To do that, you'll first need Census data to get the population of children under 5 for each county. Which county has the biggest rate of children per facility (i.e. they could use more facilities)?

```{r}
children_under_5 <- get_acs(geography = "county",
              variables = c(population = "B06001_002"),
              state = "MD",
              year = 2022)



```

```{r}
# Rename the column name 'estimate', which has our population data, with something more useful, like 'children_pop'
children_under_5 <- children_under_5 |>
  rename(children_pop = estimate)
```

```{r}
# Join the dataframes
md_childcare <- children_under_5 |>
  left_join(childcare_with_counties, join_by (GEOID == fips))


# Calculate the rates for children per facility
md_childcare <- md_childcare |>
    mutate(children_per_facility = (children_pop/total))

```

##### **A3: Once again, according to my md_childcare dataset, Prince George's county has the highest percentage of children per facility.** 

------------------------------------------------------------------------

#### **Q4** Make a map of the results for question 3. Are there any areas of the state that are more lacking in childcare facilities?

```{r}
library(sf)
md_childcare <- st_as_sf(md_childcare)
```

```{r}
ggplot(data = md_childcare) +
  geom_sf(aes(fill = children_per_facility)) +
  theme_minimal() +
  scale_fill_viridis_b(option = "magma", trans = "log") +
  labs(
    title = "Children per Childcare Facility in Maryland",
    fill = "Children per Facility"
  )


```

##### **A4**: The Eastern Shore appears to be the most lacking in childcare facilities, as evidenced by the lighter color gradients on the map. Targeted initiatives or investments in these areas could help address the imbalance.

------------------------------------------------------------------------

#### **Q5** Now let's explore if we can find patterns in under-served areas, particularly when it comes to median income of the population. First, join the childcare data median income data from the Census. Then, create a scatterplot and perform a cor.test(), and answer: is there is any correlation between income and rate of children per facilities? What makes you conclude that?

```{r}
# Again, grab data from the ACS. The ACS variable for median income is B19013_001. We want 2022 data again.
median_income <- get_acs(
  geography = "county",
  variables = c(median_income = "B19013_001"),
  state = "MD",
  year = 2022
)

```

```{r}
median_income <- median_income |>
  rename(median_income = estimate)
# Rename the column name 'estimate', which has our income data, with something more useful, like 'median_income'

```

```{r}
# Join the dataframes
# Join median income data with childcare data
md_childcare <- md_childcare |>
  left_join(median_income, by = "GEOID")


```

```{r}
ggplot(data = md_childcare, aes(x = median_income, y = children_per_facility)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Relationship Between Median Income and Children per Facility",
    x = "Median Income (USD)",
    y = "Children per Facility",
    caption = "Source: Census ACS and Childcare Data"
  )

# Create a scatterplot with median income and rate of children per facility

```

```{r}
# Perform a correlation test
cor_test_result <- cor.test(md_childcare$median_income, md_childcare$children_per_facility, use = "complete.obs")

# Print correlation test result
print(cor_test_result)
# Create a correlation test.

```

##### A5: There is no statistically significant correlation between median income and the rate of children per facility. The weak correlation and high p-value indicate that income levels are not strongly associated with the distribution of childcare facilities in Maryland counties.

------------------------------------------------------------------------

#### **Q6** Finally, let's do some text analysis. We another dataset, with inspection reports from informal child care providers in the state. This dataset has an informative column of initial findings: a column that describes issues found in these providers, if any. Your goal is to create bigrams for this column and find common issues that are mentioned in these findings. And then answer: what kind of issues listed there you think are interesting? They don't have the most frequent.

```{r}
# Read the reports data
reports <- read_csv('data/childcare-informal-provider-inspections.csv') |> clean_names()
```

```{r}
unique_words <- reports |> 
 select(initial_findings) |>
  unnest_tokens(word, initial_findings) 
View(unique_words)

data("stop_words")

stop_words <- stop_words |> 
  add_row(word = "with") |> 
  add_row(word = "for") |> 
  add_row(word = "ertg") |>
  add_row(word = "h") |>
  add_row(word = "s") |>
  add_row(word = 'in') |> 
  add_row(word = 'ecma') |> 
  add_row(word = 'standards')  
# Create a stopword dictionary (feel free to include more words)

```

```{r}
bigrams <- reports|>
  unnest_tokens(bigram, initial_findings, token = "ngrams", n = 2) |>
  separate(bigram, into = c("word1", "word2"), sep = " ")  


bigram_counts <- bigrams |>
  count(word1, word2, sort = TRUE)  

head(bigram_counts)


```

##### A6: The issues of **compliance** and **preparedness** seem noteworthy. They suggest that many informal childcare providers may struggle with meeting established standards and ensuring readiness for emergencies.These findings could reflect broader systemic challenges, such as lack of access to resources, inadequate training, or limited awareness of requirements.
