---
title: "Reverse Engineering Project"
author: "Tiasia Saunders"
date: "17 October 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
fatal_police_shootings_agencies <- read_csv("data/fatal-police-shootings-agencies.csv") 
fatal_police_shootings_data <-read_csv("data/fatal-police-shootings-data.csv") 

# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: Of the 247 women fatally shot, 48 were Black and seven of those were unarmed.
-   **Analysis summary**: We were able to get the data used in the Washington Post's article; however, it seems it has been updated since the article was published in 2020. So, we found that there have been a total of 452 women shot by police (the data was updated two days ago). Additionally, we found that there have been 85 Black females show by the police.

```{r}
fatal_police_shootings_data |> 
  group_by(gender, race) |> 
summarize(count_gender = n())

#how to override the group_by function to groups, so our output can show on the first table? 

```

### Sentence 2

-   **Sentence text**: The average age of women and men killed by police was 37, and about one-third were 25 to 34.
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{#just practicing}
{r}




# Put code to reverse engineer sentence here

# Display results of code below this codeblock



```

### Sentence 3

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
fatal_police_shootings_data |>
  filter(str_detect(armed,"unarmed")_ |>
  group_by(armed, age, gender, race) |> 
  summarise(
    
    total = n{} |>
  ))
arrange(desc(total))
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
