---
title: "Reverse Engineering Project"
author: "Tiasia Saunders"
date: "17 October 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
fatal_police_shootings_agencies <- read_csv("data/fatal-police-shootings-agencies.csv") 
fatal_police_shootings_data <-read_csv("data/fatal-police-shootings-data.csv") 

# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: Of the 247 women fatally shot, 48 were Black and seven of those were unarmed.
-   **Analysis summary**: We were able to get the data used in the Washington Post's article; however, it seems it has been updated since the article was published in 2020. So, we found that there have been a total of 452 women shot by police (the data was updated two days ago). Additionally, we found that there have been 85 Black females show by the police.

```{r}
fatal_police_shootings_data |> 
  group_by(gender, race) |> 
summarize(count_gender = n())

#how to override the group_by function to groups, so our output can show on the first table? 

```

### Sentence 2

-   **Sentence text**: About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.
-   **Analysis summary**: Since the data we are working with has been updated, we were able to find tht there is a total of 313 woman shot by police and 7,839 men shot by police. There have been 139 women shot with mental illnesses and 1,859 men shot with mental illnesses. The percentage for women would be 44% and for men 23%.

```{r}
fatal_police_shootings_data |>
  group_by(was_mental_illness_related, gender) |>
  summarise(total = n()) 
  
  total_shootings <- fatal_police_shootings_data |>
  group_by(gender) |>
  summarise(total = n())

total_shootings_mi <-fatal_police_shootings_data |>
  filter(was_mental_illness_related == "True") |>
  group_by(gender)
  summarise(total_mi = n())




# Put code to reverse engineer sentence here

# Display results of code below this codeblock



```

### Sentence 3

-   **Sentence text**: Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown. 

-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
fatal_police_shootings_data |>
  filter(str_detect(unarmed)) |>
  group_by(armed, gender, race) |> 
  summarize(
    
    total = n()) |>

arrange(desc(total))
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
