---
title: "lab_01"
author: "Derek Willis"
adapted by: "Daniel Trielli"
date: "2024-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

### **Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first. Which one is the most surprising/interesting to you?

```{r}
umd_courses |>
  group_by(department) |>
  summarise(
    mean_seats = mean(seats)

  ) |>
  arrange(desc(mean_seats))

```

**A1:** Engineering, Cooperative Education

------------------------------------------------------------------------

### **Q2** Find out the greatest number of times any individual class has been offered in our dataset. You must show the greatest number of classes first.

```{r}
umd_courses |>
  group_by(id) |>
  summarise(
    count_classes = n()
  ) |>
  arrange(desc(count_classes))

```

**A2:AASP100**

------------------------------------------------------------------------

### **Q3** Count the number of courses for each instructor and department. Which instructor has taught the most?

```{r}
umd_courses |>
  group_by(instructors, department) |>
  summarise(
    count_classes = n()
  ) |>
  arrange(desc(count_classes))
  arrange(desc(instructors))

```

**A3: Keith Yearwood- 61**

------------------------------------------------------------------------

### **Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know *how* to accomplish this; I want you to describe in words what would make the answer easier to see.

```{r}
  

```

**A4:** If the classses that didn't have professors were omitted. Such as creating a pivot table to only show the classes where classes had professor names.

------------------------------------------------------------------------

### **Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting?

1.  Which N/A course has the most instructors?

```{r}
umd_courses |>
  group_by(instructors, title) |>
  summarise(
    count_instructors = n()
  ) |>
  arrange(desc(count_instructors))




```

A: Doctoral Dissertation Research, 1176

So, I attempted to find the NA course with the most instructors. First, I grouped my table with instructors and title for courses. Then I created the column count_instructors. Lastly, I arranged that column in descending order, so I could know which N/A course had the most instructors.

How many class titles have masters in it?

1444 classes have the word master's in it.

```{r}
master_courses <-umd_courses |> 
  filter(str_detect(`title`, "Master's")) |>
  summarise(count_titles = n())|>
print(master_courses)
```

How many courses have seats over 100?

```{r}
seats_over_100 <-umd_courses |> 
  filter(seats >100) |> 
  summarise(count_seats = n())|>
  print(seats_over_100)

```
